---
title: "StatOrdPattHxC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{StatOrdPattHxC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 700, dev = "svg",
  fig.width=7, fig.height=4,
  warning = FALSE, message = FALSE
)
```

```{r setup, echo=FALSE}
library(ggplot2)
library(ggthemes)
library(reshape2)
library(tibble)
library(dplyr)
library(prodlim)
library(StatOrdPattHxC)
```

# Introduction

This package computes statistics of features from Ordinal Patterns.
Currently, it provides the asymptotic variance of the Shannon, Tsallis, and Rényi entropies, and the Fisher information measure.
With these variances, the user can apply test to check if two time series originated from similar dynamics
The package will evolve to produce other statistical characteristics.

The figure below illustrates the two main workflows.

* The user can compute the ordinal patterns, then the entropy (or Fisher information measure), and then the asymptotic variance of the entropy, or
* Can input two time series, and obtain the $p$-value of the test that verifies the hypothesis that their entropies are comparable.

![Overview of functions](./Overview.pdf){width=100%}


# Finding patterns

Let us produce three time series: two from the same dynamics (white noise smoothed with a moving averages window of size $31$, shown in green and blue), and a different one (standard normally-distributed white noise, in red).

```{r ThreeTimeSeries, out.width="90%", echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind="Mersenne-Twister")

x1 <- mov.av(rnorm(5000), order = 31) # smoothed white noise with moving average
x2 <- mov.av(rnorm(5000), order = 31) # smoothed white noise with moving average
wn <- rnorm(4970) # white noise, same size as the vectors above

time.series <- reshape2::melt(data.frame(t=1:4970, wn, x1, x2), 
                              measure.vars = 2:4)

ggplot(subset(time.series, t>1000 & t<1300), aes(x=t, y=value, col=variable)) +
  geom_line() +
  ggtitle("300 observations of three time series") +
  xlab("Time") +
  ylab("Observations") +
  theme_tufte() +
  theme(legend.position = "none")
```

We compute the sequences of ordinal patterns, and from them the histogram of patterns.
This can be done in a single step.
The following figure shows the histograms.

```{r OPsHists, out.width="90%"}

OPprobs.wn <- OPprob(wn, emb=4)
OPprobs.x1 <- OPprob(x1, emb=4)
OPprobs.x2 <- OPprob(x2, emb=4)

OPprobabilities <- reshape2::melt(data.frame(index=1:factorial(4), 
                                             OPprobs.wn, OPprobs.x1, OPprobs.x2),
                                  measure.vars = 2:4)

ggplot(OPprobabilities, aes(x=index, y=value)) +
  geom_col() +
  xlab("Pattern") +
  ylab("Proportion of patterns") +
  facet_wrap(~variable) +
  theme_tufte()
```

# Computing entropies

This package currently has three entropies and the Fisher information measure.
The input for computing all flavours of entropies, as well as the Fisher information measure, is a histogram.
We will use the histograms computed in the previous step.

```{r Entropies}
# White noise
HShannon(OPprobs.wn)
HTsallis(OPprobs.wn, beta=2)
HRenyi(OPprobs.wn, beta=1.8)
HFisher(OPprobs.wn)

# Moving averages
HShannon(OPprobs.x1)
HTsallis(OPprobs.x1, beta=2)
HRenyi(OPprobs.x1, beta=1.8)
HFisher(OPprobs.x1)
```


# Computing variances

# Applying tests

# Visualisation

The package provides the lower and upper bounds of points in the Entropy-Complexity plane
for embedding dimensions between $3$ and $6$.

```{r Boundaries, out.width="100%"}
data("LinfLsup")

ggplot(subset(LinfLsup, Side=="Lower"), 
       aes(x=H, y=C, col=Dimension, group=Dimension)) +
  geom_line() +
  geom_line(data=subset(LinfLsup, Side=="Upper"), 
            aes(x=H, y=C, col=Dimension, group=Dimension)) +
  xlab(expression(italic(H))) +
  ylab(expression(italic(C))) +
  theme_tufte()
```

# Future improvements

* Compute the Statistical Complexity variance
* Implement critical functions in C++
* Improve checking the inputs
